---
title: "Home Appraisals"
author: "Nick Hass"
date: "4/8/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(nlme) # for gls
library(gstat) # for variogram
source("glstools-master/stdres.gls.R")
source("glstools-master/predictgls.R")
```



Which variance function should we use to describe the heteroskedasticity?
Exponential correlation function!!
Bigger homes are more variable in price than smaller homes.

Spatial Linear Model $Y \sim N(X\beta, \sigma^2R)$

Change the values of the diagonal based on a covariate (living area!!!!!!!!!!)

Questions:
- Should we make House.Style, Central.Air and factor variables? What about Garage.Cars, Full.Bath, Half.Bath, and Bedroom.AbvGr ?

```{r}
homes <- read_csv("HousingPrices.csv")

# Identify factor variables -- I found that using too many breaks the gls
homes$House.Style <- as.factor(homes$House.Style)
homes$Central.Air <- as.factor(homes$Central.Air)
# homes$Garage.Cars <- as.factor(homes$Garage.Cars)
# homes$Full.Bath <- as.factor(homes$Full.Bath)
# homes$Half.Bath <- as.factor(homes$Half.Bath)
# homes$Bedroom.AbvGr <- as.factor(homes$Bedroom.AbvGr)

# Separate into data set with and without missing values
homes_na <- homes %>% filter(is.na(Price))
homes_no_na <- homes %>% filter(!is.na(Price))
```

## EDA, outliers? transformations?

```{r}
# It looks like as living area grows, price becomes more variable.
par(mfrow = c(2,2))
ggplot(homes_no_na, mapping = aes(x = Gr.Liv.Area, y = Price)) +
  geom_point() +
  labs(title = "Price by Living Area",
       x = "Living Area (sq feet)")

# Newer homes tend to be more expensive
ggplot(homes_no_na, mapping = aes(x = Year.Remod.Add, y = Price)) +
  geom_point() +
  labs(title = "Price by Remodel Date",
       x = "Year of Remodel")

ggplot(homes_no_na, mapping = aes(x = Central.Air, y = Price)) +
  geom_boxplot() +
  labs(title = "Price by Central Air",
       x = "Central Air") +
  scale_x_discrete(labels = c("N" = "No", "Y" = "Yes"))

ggplot(homes_no_na, mapping = aes(x = House.Style, y = Price)) +
  geom_boxplot() +
  labs(title = "Price by House Style",
       x = "House Style") +
  scale_x_discrete(labels = c("1Story" = "1 Story", "2Story" = "2 Story",
                              "SLvl" = "Split Level"))
### ***update x axis labels for these plots***

# Larger garages tend to be more expensive homes
ggplot(homes_no_na, mapping = aes(x = as.factor(Garage.Cars), y = Price)) +
  geom_boxplot() +
  labs(title = "Price by Size of Garage",
       x = "Size of Garage in Car Capacity")

ggplot(homes_no_na, mapping = aes(x = as.factor(Half.Bath), y = Price)) +
  geom_boxplot() +
  labs(title = "Price by Number of Half Bathrooms",
       x = "Half Bathrooms")

ggplot(homes_no_na, mapping = aes(x = as.factor(Full.Bath), y = Price)) +
  geom_boxplot() +
  labs(title = "Price by Number of Full Bathrooms",
       x = "Full Bathrooms")

ggplot(homes_no_na, mapping = aes(x = as.factor(Bedroom.AbvGr), y = Price)) +
  geom_boxplot() +
  labs(title = "Price by Number of Bedrooms",
       x = "Bedrooms")

```


## Model Specification

MLR with Spatial Correlation Structure & Exponential Variance (Heteroskedasticity)

$$
Y \sim N(X\beta, \Sigma)
$$

$$
\Sigma = DRD
$$

$$
D = diag(\sigma_1, ..., \sigma_n)
$$

$$
R = \text{Correlation Matrix}
$$

## Model Fitting

Simple Model
```{r}
homes.lm <- lm(Price ~ ., data = homes_no_na)
BIC(homes.lm)

myVariogram <- variogram(object = Price ~ Gr.Liv.Area + House.Style + 
                                  Year.Remod.Add + Central.Air + Full.Bath + 
                                  Half.Bath + Bedroom.AbvGr + Garage.Cars,
                         locations = ~Lon + Lat, 
                         data = homes_no_na)

plot(myVariogram, main = "Variogram of the Residuals") # There is spatial correlation.
```


Correlation
```{r}
homes.lm.hetero <- gls(model = Price ~ Gr.Liv.Area + House.Style + 
                               Year.Remod.Add + Central.Air + Full.Bath + 
                               Half.Bath + Bedroom.AbvGr + Garage.Cars, 
    data=homes_no_na,  # data without missing values
    weights = varExp(form=~Gr.Liv.Area), # D part (Heteroskedastic)
    correlation = corExp(form = ~Lat+Lon, nugget = TRUE), # R part (Spatial corr)
    method="ML") ## with factor central air and house style

AIC(homes.lm.hetero)
```

## Model Validation

```{r, fig.cap = "Model Validation Plots", figures-side, fig.show="hold", out.width="50%"}
# Spatial MLR Model with heteroskedasticity: homes.lm.hetero

# Linearity: avPlots
car::avPlots(homes.lm, main = "Check for Linearity")

# Independence by decorrelating residuals and looking at the variogram of decorrelated residuals.
sres <- stdres.gls(homes.lm.hetero) # Decorrelate residuals
residDF <- data.frame(Lon=homes_no_na$Lon, Lat=homes_no_na$Lat, decorrResid=sres)
residVariogram <- variogram(object=decorrResid~1, locations=~Lon+Lat, data=residDF)
plot(residVariogram, main = "Variogram of the Standardized Decorrelated Residuals")

#  Normality: Histogram of standardized decorrelated residuals
hist(sres, probability = TRUE, ylim = c(0, 0.4), xlim = c(-5, 5),
     main = "Standardized Decorrelated Residuals")
curve(dnorm(x), add = TRUE, col = "blue") # Normality holds

# Equal Variance: Fitted values vs. Standardized and decorrelated residuals plot
fitted_values <- fitted(homes.lm.hetero)
ggplot(homes_no_na, aes(x = fitted_values, y = sres)) +
  geom_point() + ggtitle("Equal Variance Check") # Equal variance holds
```
As shown by the added-variable plots, linearity holds.

After accounting for spatial correlation by using a spatial correlation function, independence holds, as shown by the variogram of the standardized and decorrelated residuals.

As shown by the histogram of standardized and decorrelated residuals, normality holds.

After accounting for heteroskedacity by using a exponential variance function, equal variance holds, as shown by the fitted values by standardized and decorrelated residuals plot.

## Statistical Inference
```{r}

```

Steps:
EDA
Model Specification
Model Fitting
Model Validation
Statistical Inference


Q1: How well do the home characteristics explain sale price?
```{r}
cor(homes_no_na$Price, fitted_values)^2  # Pseudo R squared = 0.9329216
```

A1: Home characteristics explain 93.29216% of the variability in home prices.

Q2: What factors increase the sale price of a home?
```{r}
homes.lm.hetero$coefficients > 0
```

A2: Above ground living area, Split level homes, newer remodels, having central air, more half bathrooms, and larger sized garages increase the sale price of a home.

Q3: Does the variability of sale price increase with the size of the home (as given by living area)?
```{r}
# This plot is also seen in the EDA
ggplot(homes_no_na, mapping = aes(x = Gr.Liv.Area, y = Price)) +
  geom_point() +
  labs(title = "Price by Living Area")
```

A3: Yes, as can be seen from this histogram, as living area increases, there is a larger range of home prices.

Q4: What is your predicted/appraised sale price for the homes in the dataset that do not have a saleprice? Include your predictions in a map
```{r}
missingPreds <- predictgls(homes.lm.hetero, newdframe = homes_na, level = 0.95)  # predict for missing values, save df
not_misssing_preds <- predictgls(homes.lm.hetero, newdframe = homes_no_na)

homes_na$Price <- missingPreds$Prediction

full <- rbind(homes_no_na, homes_na)  # full <- rbind(missing, without missing)


ggplot(data = homes, aes(x = Lon, y = Lat)) +
  geom_point(aes(col = Price), size = 2.5) +
  scale_color_distiller(palette = "Spectral", na.value = NA) +
  labs(title = "Home Price Data") +
  theme_bw()

# ggplot(data = full, aes(x = Lon, y = Lat)) + 
#   geom_point(aes(col = Price), size = 2.5) +
#   scale_color_distiller(palette = "Spectral", na.value = NA) +
#   labs(title = "Home Price Predictions")
options(scipen=999)
#options(scipen=0, digits=7)
ggplot(mapping = aes(x = Lon, y = Lat)) + 
  geom_point(data = full[1:465,], aes(col = Price), size = 2.5, alpha = 0.5) +
  geom_point(data = full[466:517,], aes(col = Price), size = 2.5, shape = 13,
             alpha = 1) +
  theme_dark() +
  scale_color_distiller(palette = "Spectral", na.value = NA) +
  labs(title = "Home Price Predictions with Original Data",
       subtitle = "Predictions Marked with Target Shaped Points") +
  scale_fill_continuous(labels=scales::comma)

```
